---
title: "R Notebook"
output: html_notebook
---

```{r}
library(readxl)
library(tidyverse)
```

Zaczynamy od wczytnia danych

```{r}
dane <- read_excel("../data/test.xlsx")
```

## Exercise 1. 

1	single
2	married
3	widow
4	divorced 
5	legal separation
6	actual separation

```{r}
table(dane$fc11)
```

Odpowiedź: 1062 married persons took part in the survey (variable fc11).

# Exercise 2. 

The average age (variable age) of married persons (variable fc11) was ...........years

```{r}
dane_zwiazek <- dane %>% filter(fc11 == 2)
mean(dane_zwiazek$age)
```

```{r}
dane %>%
  filter(fc11 == 2) %>%
  summarise(m = mean(age))
```

```{r}
aggregate(age ~ fc11, data = dane, FUN = mean)
```

```{r}
mean(dane$age[dane$fc11 == 2])
```


The average age (variable age) of married persons (variable fc11) was 44.8 years

# Exercise 3. 


```{r}
aggregate(fp65 ~ sex, data = dane, FUN = min)
```

```{r}
dane %>%
  filter(sex == 1) %>%
  summarise(min = min(fp65),
            max = max(fp65),
            mean = mean(fp65),
            sd = sd(fp65),
            median = median(fp65))
```

Compute basic descripive statistics of income (variable fp65) (min, max, mean, standard deviaKon and median) for males (variable sex). [2p]

Min=200
Max= 30000
Mean= 2251.423
Sd= 1898.573
Median=2000


# Exercise 4. 

Complete the following conKngency table, which contains informaKon about the number of respondents by class of locality (variable class) and the variable fp29 (What is more important in your life?).

```{r}
xtabs(~ class + fp29, dat = dane)
```

```{r}
table(dane$class, dane$fp29)
```

```{r}
dane %>%
  count(class, fp29) %>%
  spread(fp29, n)
```

# Exercise 5.

Recode variable age into two levels: 1 – up to 35 years (including 35), 2 – above 35 years and fill in the table below with frequencies. [2p]


```{r}
dane$wiek <- ifelse(dane$age <= 35, 1, 2)
xtabs(~ sex + wiek, dane)
```

# Exercise 6. 

Complete informaKon about the percentage of respondents by marital status (variable fc11) by entering the appropriate values above each bar separately.

```{r}
table(dane$fc11) / nrow(dane) * 100
```

```{r}
prop.table(table(dane$fc11))*100
```

```{r}
dane %>%
  count(fc11) %>%
  mutate(p = n/sum(n)*100)
```

# Exercise 7. 

Compute and interpret the Pearson coefficient of correlaKon between height (variable fp55) and weight (variable fp56) for males (variable sex) living in 500T+ ciKes (variable class).

```{r}
mez_city <- dane %>% filter(sex == 1, class == 1)
cor(mez_city$fp55, mez_city$fp56)
```

```{r}
dane %>% 
  filter(sex == 1, class == 1) %>% 
  summarise(cor = cor(fp55, fp56))
```

```{r}
cor(dane[dane$sex ==1 & dane$class == 1, c("fp55", "fp56")], method = "pearson")
```

# Exercise 8. 

QuarKles of height (variable fp55) for females (variable sex) are equal to: [2p] 

Q1=...................... Q2=...................... Q3=......................


```{r}
quantile(x = dane$fp55[dane$sex == 2])
```

```{r}
dane %>%
  filter(sex == 2 ) %>%
  summarise(q1 = quantile(fp55, probs = 0.25),
            q2 = quantile(fp55, probs = 0.50),
            q3 = quantile(fp55, probs = 0.75))
```


# Exercise 9. 

```{r}
rysuj_wykres <- function(zmienna) {
  barplot(table(zmienna))
}

rysuj_wykres(dane$class)
```

# Exercise 10. 

Write an R funcKon which for a given radius r returns the surface area of the ball


```{r}
licz_pow_i_obj <- function(r) {
  
  wynik <- c(pow = 4*pi*r^2, obj = 4/3*pi*r^3)
  
  return(wynik)
}
```

```{r}
licz_pow_i_obj(10)
```

